Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-02-10 constantRate/ConstCompFuzzExt.tex
adding notes from people I had read

\begin{abstract}
Fuzzy extractors convert a noisy source of entropy into a consistent uniformly distributed key.  In the process of eliminating noise, they  lose some of the entropy of the original source---in the worst case, as much as the logarithm of the number of correctable error patterns. We call what is left after this worst-case loss the \emph{minimum usable entropy}.  Unfortunately, this quantity is negative for some sources that are important in practice. Most known approaches for building fuzzy extractors work in the worst case and cannot be used when the minimum usable entropy is negative.

@@

Our first construction requires high individual entropy of a constant fraction of symbols, but permits symbols to be dependent.  Our second construction requires  a constant fraction of symbols to have a constant amount of entropy conditioned on prior symbols. The constructions can be implemented efficiently based on number-theoretic assumptions or assumptions on cryptographic hash functions.
\end{abstract}

@@

\paragraph{Limitations of Entropy-Based Approaches}
Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess some $w'$ within distance $t$ of $w$ they will be able to easily distinguish $r$ from uniform by running $\rep$. If $t$ is larger, then $\rep$ tolerates more $w'$ values, so the adversary's job is easier. In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then the \emph{entire distribution} of $w$ may have min-entropy $m$ and be contained in a single ball.  Then the adversary can always produce the key by running $\rep$ with the center of the ball as $w'$.

More generally, let $B_t$ denote the number of points in a ball of radius $t$. For any $m$ and $t$, there is a distribution of min-entropy $m$ such that the adversary can guess a correct $w'$ with probability $1/\lceil 2^m/B_t) \rceil\approx B_t 2^{-m}$: the distribution consists of the uniform distribution over all points in several nonoverlapping balls of radius $t$ (the metric space must be large enough for these balls not to intersect). We thus call $m-\log B_t$ the \emph{minimum usable} entropy, denoted by $\Huse$. The previous paragraph shows that  no fuzzy extractor can handle all distributions of a given min-etropy $m$ if  $\Huse\le 0$.

Prime candidate sources for authentication have $\Huse\le 0$.  As an example, the iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  The best estimate for iris entropy is $249$ bits~\cite{daugman2004}.  Daugman uses specialized wavelets to derive a $2048$ bit string called an iris code.  Let the outcome of this transform~(on different irises) define a distribution $w$.  The precise number of errors that must be tolerated depends on the desired false reject rate (how often the correct key is produced).  For a false reject rate of $\le 80\%$, a $t$ of approximately $205$ is required.  We have the following calculation for $\Huse$:
\[
\Huse = \Hoo(W) - \log |B_t|
= 249 - \log \sum_{i=0}^{205} {2048 \choose i} \approx -707.
\]
There is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining} to Daugman but it does not affect $\Huse$ dramatically.\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but underestimates the number of possible error patterns; their calculation, which we cannot confirm, is $\Huse \approx 44$.}

\paragraph{Our Contributions}
We provide the first constructions of computational fuzzy extractors that can be used for a large class of distributions with $\Huse\le 0$ over $\mathcal{Z}^\ell$ for a large alphabet $\mathcal{Z}$.  As explained above, such constructions cannot work without some restriction on the distribution.
Our first construction is secure when symbols in $w$
each have individual superlogarithmic min-entropy, even if they are arbitrarily correlated. Moreover,
a constant fraction of symbols in $w$ may have little entropy, as long as knowledge of their values does not reduce the entropy of the high-entropy symbols too much (see \defref{def:block guessable}).

>>>>

\begin{abstract}
Fuzzy extractors convert a noisy source of entropy into a consistent uniformly-distributed key.  In the process of eliminating noise, they  lose some of the entropy of the original source---in the worst case, as much as the logarithm of the number of correctable error patterns. We call what is left after this worst-case loss the \emph{minimum usable entropy}.  Unfortunately, this quantity is negative for some sources that are important in practice. Most known approaches for building fuzzy extractors work in the worst case and cannot be used when the minimum usable entropy is negative.

@@

Our first construction requires high individual entropy of a constant fraction of symbols, but permits symbols to be dependent.  Our second construction requires  a constant fraction of symbols to have a constant amount of entropy conditioned on prior symbols. The constructions can be implemented efficiently based on number-theoretic assumptions or assumptions on cryptographic hash functions.
\bnote{more about what we do}
\end{abstract}

@@

\paragraph{Limitations of Entropy-Based Approaches}
\bnote{shorten this heading?}
Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess some $w'$ within distance $t$ of $w$ they will be able to easily distinguish $r$ from uniform by running $\rep$. If $t$ is larger, then $\rep$ tolerates more $w'$ values, so the adversary's job is easier. In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then the \emph{entire distribution} of $w$ may have min-entropy $m$ and be contained in a single ball.  Then the adversary can always produce the key by running $\rep$ with the center of the ball as $w'$.\bnote{was told this wasn't clear enough}

More generally, let $B_t$ denote the number of points in a ball of radius $t$. For any $m$ and $t$, there is a distribution of min-entropy $m$ such that the adversary can guess a correct $w'$ with probability $1/\lceil( 2^m/B_t) \rceil\approx B_t 2^{-m}$: the distribution consists of the uniform distribution over all points in several non-overlapping balls of radius $t$ (the metric space must be large enough for these balls not to intersect). We thus call $m-\log B_t$ the \emph{minimum usable} entropy, denoted by $\Huse$. The previous paragraph shows that  no fuzzy extractor can handle all distributions of a given min-entropy $m$ if  $\Huse\le 0$.
\bnote{is there anyway we can reasonably say what previous works support?}

Prime candidate sources for authentication have $\Huse\le 0$.  As an example, the iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  Daugman estimates irises contain $249$ bits of entropy~\cite{daugman2004}.  Daugman uses specialized wavelets to derive a $2048$ bit string called an iris code.  Let the outcome of this transform~(on different irises) define a distribution $w$.  The precise number of errors that must be tolerated depends on the desired false reject rate (how often the correct key is produced).  For a false reject rate of $\le 80\%$, a $t$ of approximately $205$ is required.  We have the following calculation for $\Huse$:
\[
\Huse = \Hoo(W) - \log |B_t|
= 249 - \log \sum_{i=0}^{205} {2048 \choose i} \approx -707.
\]
There is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining} to Daugman but it does not affect $\Huse$ dramatically.\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but their analysis underestimates the number of possible error patterns; their calculation, which we cannot confirm, is $\Huse \approx 44$.}

\paragraph{Our Contributions}
We provide the first constructions of computational fuzzy extractors that can be used for a large class of distributions with $\Huse\le 0$ over $\mathcal{Z}^\ell$ for a large alphabet $\mathcal{Z}$.  As explained above, such constructions cannot work without some restriction on the distribution.
Our first construction is secure when symbols in $w$
each have individual super-logarithmic min-entropy, even if they are arbitrarily correlated. Moreover,
a constant fraction of symbols in $w$ may have little entropy, as long as knowledge of their values does not reduce the entropy of the high-entropy symbols too much (see \defref{def:block guessable}).
